Accelerated Inference for Partially Observed Stochastic Processes with Automatic Differentiation

### 2024-05-30 paper/ms.tex
Updates from Overleaf

\subsection{MOP-$\alpha$ Bias and Variance}
\begin{thm}
    \label{thm:mop-biasvar}
    When $\alpha\in(0,1)$, $\theta=\phi$, the MSE and variance of MOP-$\alpha$, where $\psi(\alpha)=(\alpha^k  + \alpha^{k+1} - \alpha)(1-\alpha)^{-1}$, are:
    \vspace*{-1ex}
    \begin{align*}
        &\E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_2^2 
        \\
        &\lesssim \min_{k \leq N} NpG'(\theta)^2\left(k^2J^{-1}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\psi(\alpha)\right),\\[-1ex]
        &\text{Var}(\nabla_\theta \hat\ell^{\alpha}(\theta)) \lesssim \min_{k\leq N} NpG'(\theta)^2\left(\frac{k^2}{(1-\alpha)^2J} + \frac{16\alpha^{2k}}{(1-\alpha)^2}N\right).
        \end{align*}
\end{thm}
\kevin{Explain exactly what $\epsilon$ is here, and interpret the rates as agreeing with our intuition. Talk about the relationship between the theory and the empirical results in the bias-variance tradeoff figure above, as well as the phase transition between $Nk^2/J$ and $N^4/J$ variance with MOP-$\alpha$ and MOP-$1$, where $k$ is really a surrogate for $\alpha$ -- with more memory, you truncate less. }

>>>>

\subsection{MOP-$\alpha$ Bias and Variance}
\begin{thm}
    \label{thm:mop-biasvar}
    When $\alpha\in(0,1)$, $\theta=\phi$, the MSE and variance of MOP-$\alpha$, where $\psi(\alpha)=(\alpha^k  + \alpha^{k+1} - \alpha)(1-\alpha)^{-1}$ and $\epsilon$ depends on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ as in \cite{karlajainen2001} are:
    \vspace*{-1ex}
    \begin{align*}
        &\E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_2^2 
        \\
        &\lesssim \min_{k \leq N} NpG'(\theta)^2\left(k^2J^{-1}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\psi(\alpha)\right),\\[-1ex]
        &\text{Var}(\nabla_\theta \hat\ell^{\alpha}(\theta)) \lesssim \min_{k\leq N} NpG'(\theta)^2\left(\frac{k^2}{(1-\alpha)^2J} + \frac{16\alpha^{2k}}{(1-\alpha)^2}N\right).
        \end{align*}
\end{thm}
\kevin{Explain exactly what $\epsilon$ is here, and interpret the rates as agreeing with our intuition. Talk about the relationship between the theory and the empirical results in the bias-variance tradeoff figure above, as well as the phase transition between $Nk^2/J$ and $N^4/J$ variance with MOP-$\alpha$ and MOP-$1$, where $k$ is really a surrogate for $\alpha$ -- with more memory, you truncate less. }
